Tempus: Probabilistic Network Latency Verifier

### 2022-10-17 paper/main.tex
[Paper] Added 9 paragraphs of section 3

>>>>

\section{Overview}
To be written

\section{On Latency Modeling}
\subsection{Latency as Path Property} 
We began our study by first pondering about a basic construct: modeling the end-to-end latency of 
a single packet.
In a packet-switched network, a packet is sent from one transmiting host to one receiving host 
through the many components of the network (e.g. links, routers, firewall) and each of those 
components might introduce some latency into the packet transmission process.
Figuring out which of those components will actually introduce latency into the packet in question, 
and by how much, is the next logical step that we must figure out.

Obviously, a given packet doesn't need to visit all network components to reach its destination 
host.
A network engineer will configure the network in such a way that a packet will only need to be 
routed via a specific subset of its components.
That specific subset is dictated by two things: how those components are connected together 
(i.e. topology) and how the control plane protocols are configured to route a packet (e.g. routing 
protocol, ACL).
Based on a variety of these setups, nodes in the network will form a forwarding table to route a 
given packet appropriately.

The goal of a classical control plane verifier then, is to use this forwarding table in some 
form to verify certain properties. %TODO: examples?
However, looking at the forwarding table alone will not give us a conclusive result regarding 
which components are going to be visited by a given packet, making verification of latency 
properties less clear.
For example, the network might be configured with a load-balancing protocol in which a packet 
departing from a source host might take multiple different \textit{paths} (with certain 
probability weights) to arrive at the destination host, possibly resulting in a different 
end-to-end latency measurement.

Therefore, we argue that when it comes to analyzing end-to-end latency, a network path should be 
the primary unit of reasoning, rather than forwarding table.
By being more specific about our unit of reasoning, we could answer verification questions more 
clearly, and we could design our verifier more efficiently since we could use it to represent 
multiple different forwarding tables that shares the same path.

\subsection{Relation to Classical Verifier}
Before we analyze the latency of a given packet that propagates through a certain network 
path however, we must make sure that that path exists in the first place.
We note that latency is a property that only make sense after connectivity between two hosts 
has been established. 
In other words, if two nodes in a network aren't even functionally connected (e.g. physical 
link failure, ACL policy), then the latency between them will \textit{always} be infinite, 
making the verification task trivial.

Fortunately, there are a rich body of work in the network verification literature regarding 
functional reachability under failure. %TODO: cite
We could then design our verifier on top of an existing classical control plane verifier.
We use it to verify reachability property, and only if the reachability property is fulfilled, 
we would verify whether the latency between two hosts fulfilled some additional condition.

\subsection{Latency Distribution}
Up to this point, we have talked about measuring the latency of a single packet by figuring out 
its path; analyzing which exact components it has traversed through.
However, when we try to generalize this framework and ask questions about the latency of multiple 
packets, it is apparent that the path alone is not a determinative information, as the latency 
of two packets propagating through the same path might be different due to a multitude of factors.

The natural extension to the framework then, is instead of representing latency of a path with a 
single number, we instead represent it with a continuous random variable that signifies the 
possible delay that a given packet traversing through that path might have.
This random variable will have a distribution that marginalizes over all other factors other than 
the path.

We can then use this latency distribution to verify some temporal properties in a probabilistic 
manner.
For example, we could verify the probability that a packet will be delivered in under a time 
unit by taking the CDF of the distribution.

\subsection{Path Decomposition and Convolution}
Expand on these points:
\begin{itemize}
    \item How do we get the path latency distribution? Component-wise measurement data
    \item What measurement? link delay and queuing delay $\rightarrow$ not all delays are 
    significant
    \item What to do with it? Simulation $\rightarrow$ lack of error guarantee
    \item Convolution
    \item Independence?
\end{itemize}
